---
title: Clustering with R
author: Michael Bach
date: '2020-10-25'
slug: clustering-with-r
categories: ["data-science"]
images: ~
---

---------------------------------------

> **TL;DR**
>
> Some summary.

---------------------------------------

Clustering is an unsupervised learning method and helps grouping data points into clusters based on some similarity measures. As there are many clustering techniques out there, it's vital to make sure to understand the underlying task and variables to choose the appropriate technique. Some common methods are:

- Partitional Clustering
- Hierarchical Clustring
- Fuzzy Clustering
- Density-Based Clustering
- Model-Based Clustering

In this post I'll cover the ones I've used the most so far. I'll give some examples using the popular *iris* data set.

## The data set

```{r setup, include=FALSE}
options(scipen = 10)
library(data.table)
library(plotly)
dt <- iris
setDT(dt)
colnames(dt) <- tolower(colnames(dt))
colnames(dt) <- gsub("\\.", "_", colnames(dt)) # I prefer snake_case, but just a matter of taste
```

The data set consists of 3 species, encoded as factors (*setosa*, *versicolor*, *virginica*). These species are all flowers and have different sepal and petal length and width, which are reported.

If we want to simply plot the data by group, we'll see that the groups differ nicely from each other.

```{r plot_species}
a <- list(text = "Sepal Width", x = 1, y = 8.5, allign = "center", showarrow = FALSE)
fig_sepal_w <- plot_ly(data = dt, y = ~sepal_width, color = ~species, type = "box") %>%
  layout(annotations = a)

a$text <- "Sepal Length"
fig_sepal_l <- plot_ly(data = dt, y = ~sepal_length, color = ~species, type = "box") %>% 
  layout(annotations = a)

a$text <- "Petal Width"
fig_petal_w <- plot_ly(data = dt, y = ~petal_width, color = ~species, type = "box") %>% 
  layout(annotations = a)

a$text <- "Petal Length"
fig_petal_l <- plot_ly(data = dt, y = ~petal_length, color = ~species, type = "box") %>% 
  layout(annotations = a)

fig <- subplot(fig_sepal_w, fig_sepal_l, fig_petal_w, fig_petal_l, shareY = TRUE) %>% 
  layout(showlegend = FALSE)

fig
```

Based on this first intuition, we should get some nice clustering results on the data set.

## K-Means Clustering

To begin with, there is the all time favorite when it comes to clustering - the *k-means* algorithm. This algorithm basically tries to partition the observations in *k* clusters.


## Hierarchical Clustering

https://towardsdatascience.com/the-5-clustering-algorithms-data-scientists-need-to-know-a36d136ef68
https://www.datanovia.com/en/blog/types-of-clustering-methods-overview-and-quick-start-r-code/
https://www.researchgate.net/publication/220571682_An_overview_of_clustering_methods


